---
title: "Broad scan figure code | factor_corr = 0.3"
output:
  html_document:
    toc: true
    toc_depth: 2
date: "2026-06-15"
---

# Purpose

This file contains only the code for the broad-scan 2x2 figures used in the thesis.

It reproduces the two broad scans for `factor_corr = 0.3`:

- J-scan: fixed `N = 300`, varying `J = 9, 12, ..., 45`
- N-scan: fixed `J = 30`, varying `N = 100, 200, ..., 1000`

Each scan produces a 2x2 panel with MSE, average absolute bias, variance, and runtime.

# Setup

```{r setup, message=FALSE, warning=FALSE}
# Download the helper scripts and replace the example paths with your own local paths.

# Load the required helper scripts. Replace the example paths with your own local paths.
source("C:/Users/YourName/Desktop/BEP/R/rESEM-CC.R")
source("C:/Users/YourName/Desktop/BEP/R/generation_function.R")

library(MASS)
library(lavaan)
library(lslx)
```

# Helper functions

```{r helper_functions}
# Build the lslx model syntax for the correlated-factor lslx model.
build_lslx_model <- function(J, n_factors = 3) {
  J_per_factor <- J / n_factors     # Number of primary items per factor
  lines <- c()                      # Store model syntax lines

  # Add primary loadings.
  for (k in 1:n_factors) {
    main_items <- paste0("x", ((k - 1) * J_per_factor + 1):(k * J_per_factor))
    lines <- c(lines, paste0("f", k, " :=> ", paste(main_items, collapse = " + ")))
  }

  # Add penalized cross-loadings.
  for (k in 1:n_factors) {
    primary_items <- ((k - 1) * J_per_factor + 1):(k * J_per_factor)
    other_items <- paste0("x", setdiff(1:J, primary_items))
    lines <- c(lines, paste0("f", k, " :~> ", paste(other_items, collapse = " + ")))
  }

  # Fix factor variances to one.
  for (k in 1:n_factors) {
    lines <- c(lines, paste0("f", k, " <=> fix(1) * f", k))
  }

  # Factor covariances are not fixed here, matching the correlated-factor scans.

  paste(lines, collapse = "\n")
}


# Align estimated loadings to the true loading matrix.
align_loadings <- function(Lambda_est, Lambda_true) {
  n_factors <- ncol(Lambda_true)
  cor_mat <- matrix(NA, n_factors, n_factors)

  for (i in 1:n_factors) {
    for (j in 1:n_factors) {
      cor_mat[i, j] <- abs(cor(Lambda_est[, i], Lambda_true[, j]))
    }
  }

  order_est <- rep(NA, n_factors)   # Matched factor order
  used <- rep(FALSE, n_factors)     # Prevent duplicate matching

  for (j in 1:n_factors) {
    best_i <- which.max(cor_mat[, j] * (!used))
    order_est[j] <- best_i
    used[best_i] <- TRUE
  }

  Lambda_aligned <- Lambda_est[, order_est]

  for (j in 1:n_factors) {
    if (cor(Lambda_aligned[, j], Lambda_true[, j]) < 0) {
      Lambda_aligned[, j] <- -Lambda_aligned[, j]
    }
  }

  colnames(Lambda_aligned) <- paste0("f", 1:n_factors)
  Lambda_aligned
}


# Fit lslx to one simulated dataset.
run_lslx <- function(X, J, n_factors = 3, Lambda_true = NULL) {
  X_df <- as.data.frame(scale(as.matrix(X)))   # Standardize observed data
  colnames(X_df) <- paste0("x", 1:J)           # Variable names used in model syntax

  model <- build_lslx_model(J = J, n_factors = n_factors)
  lambda_grid <- seq(0.00, 0.10, 0.05)         # Candidate LASSO penalties

  result <- tryCatch({
    fit <- plsem(
      model = model,
      data = X_df,
      penalty_method = "lasso",
      lambda_grid = lambda_grid
    )

    loading_matrix <- fit$extract_coefficient_matrix(
      selector = "bic",      # Select final solution by BIC
      block = "y<-f"         # Extract loading matrix
    )$g

    rownames(loading_matrix) <- paste0("x", 1:J)
    colnames(loading_matrix) <- paste0("f", 1:n_factors)

    if (!is.null(Lambda_true)) {
      loading_matrix <- align_loadings(loading_matrix, Lambda_true)
    }

    list(converged = TRUE, loadings = loading_matrix)
  }, error = function(e) {
    list(converged = FALSE, loadings = NULL, error_msg = conditionMessage(e))
  })

  result
}


# Fit cardinality-constrained Regularized ESEM to one simulated dataset.
run_cc_esem <- function(X, J, n_factors = 3, card = NULL,
                        nstarts = 20, Lambda_true = NULL) {
  if (is.null(card)) {
    card <- rep(J / n_factors, n_factors)       # True cardinality per factor
  }

  X_scaled <- scale(as.matrix(X))               # Standardize observed data

  result <- tryCatch({
    fit <- MULTISTART_CCrESEM(
      DATA = X_scaled,
      R = n_factors,
      CARD = card,
      MaxIter = 1000,
      eps = 1e-6,
      nstarts = nstarts
    )

    loading_matrix <- fit$loadings
    rownames(loading_matrix) <- paste0("x", 1:J)
    colnames(loading_matrix) <- paste0("f", 1:n_factors)

    if (!is.null(Lambda_true)) {
      loading_matrix <- align_loadings(loading_matrix, Lambda_true)
    }

    list(converged = fit$converged, loadings = loading_matrix)
  }, error = function(e) {
    list(converged = FALSE, loadings = NULL, error_msg = conditionMessage(e))
  })

  result
}


# Summarize MSE, AB, VAR, runtime for one condition.
summarize_condition <- function(lslx_list, esem_list, lslx_time, esem_time, lslx_conv, Lambda_true) {
  lslx_valid <- Filter(Negate(is.null), lslx_list)
  esem_valid <- Filter(Negate(is.null), esem_list)

  if (length(lslx_valid) > 1) {
    lslx_arr <- simplify2array(lslx_valid)
    lslx_mean_lm <- apply(lslx_arr, c(1, 2), mean)
    lslx_mse_vec <- sapply(lslx_valid, function(lm) mean((lm - Lambda_true)^2))

    lslx_mse <- mean(lslx_mse_vec)
    lslx_mse_sd <- sd(lslx_mse_vec)
    lslx_ab <- mean(abs(lslx_mean_lm - Lambda_true))
    lslx_var <- mean(apply(lslx_arr, c(1, 2), var))
  } else {
    lslx_mse <- NA
    lslx_mse_sd <- NA
    lslx_ab <- NA
    lslx_var <- NA
  }

  if (length(esem_valid) > 1) {
    esem_arr <- simplify2array(esem_valid)
    esem_mean_lm <- apply(esem_arr, c(1, 2), mean)
    esem_mse_vec <- sapply(esem_valid, function(lm) mean((lm - Lambda_true)^2))

    esem_mse <- mean(esem_mse_vec)
    esem_mse_sd <- sd(esem_mse_vec)
    esem_ab <- mean(abs(esem_mean_lm - Lambda_true))
    esem_var <- mean(apply(esem_arr, c(1, 2), var))
  } else {
    esem_mse <- NA
    esem_mse_sd <- NA
    esem_ab <- NA
    esem_var <- NA
  }

  data.frame(
    lslx_mse_mean = lslx_mse,
    lslx_mse_sd = lslx_mse_sd,
    esem_mse_mean = esem_mse,
    esem_mse_sd = esem_mse_sd,
    lslx_ab = lslx_ab,
    esem_ab = esem_ab,
    lslx_var = lslx_var,
    esem_var = esem_var,
    lslx_time = mean(lslx_time),
    esem_time = mean(esem_time),
    lslx_conv_rate = mean(lslx_conv)
  )
}


# Run one broad scan over J or N.
run_scan <- function(scan_type = c("J", "N"),
                     J_values = NULL,
                     N_values = NULL,
                     fixed_J = NULL,
                     fixed_N = NULL,
                     factor_corr = 0.3,
                     n_rep = 50) {

  scan_type <- match.arg(scan_type)
  results <- list()

  if (scan_type == "J") {
    loop_values <- J_values
  } else {
    loop_values <- N_values
  }

  for (value in loop_values) {
    J <- if (scan_type == "J") value else fixed_J
    N <- if (scan_type == "N") value else fixed_N

    cat("Running", scan_type, "scan | J =", J, "| N =", N, "\n")

    lslx_lm_list <- vector("list", n_rep)
    esem_lm_list <- vector("list", n_rep)
    lslx_conv <- logical(n_rep)
    lslx_time_vec <- numeric(n_rep)
    esem_time_vec <- numeric(n_rep)

    for (rep in 1:n_rep) {
      dgp <- data_generation(
        PL = 0.6,                       # True primary loading
        nvar.factor = J / 3,            # Number of items per factor
        nfactors = 3,                   # Number of latent factors
        reliability = 0.36,             # Explained item variance, equal to 0.6^2
        factor_corr = factor_corr,      # Population factor correlation
        N = N,                          # Sample size
        beta = c(0.5, 0.3, 0.2),        # Factor effects on generated outcome y
        VAFy = 0.5,                     # Variance of y explained by factors
        CL = FALSE                      # No population cross-loadings
      )

      t1 <- system.time({
        r1 <- run_lslx(X = dgp$X, J = J, Lambda_true = dgp$Ptrue)
      })

      t2 <- system.time({
        r2 <- run_cc_esem(X = dgp$X, J = J, Lambda_true = dgp$Ptrue)
      })

      lslx_conv[rep] <- r1$converged
      lslx_lm_list[[rep]] <- if (r1$converged) r1$loadings else NULL
      esem_lm_list[[rep]] <- if (r2$converged) r2$loadings else NULL
      lslx_time_vec[rep] <- t1["elapsed"]
      esem_time_vec[rep] <- t2["elapsed"]
    }

    summary_row <- summarize_condition(
      lslx_list = lslx_lm_list,
      esem_list = esem_lm_list,
      lslx_time = lslx_time_vec,
      esem_time = esem_time_vec,
      lslx_conv = lslx_conv,
      Lambda_true = dgp$Ptrue
    )

    results[[as.character(value)]] <- cbind(
      data.frame(J = J, N = N, JN_ratio = round(J / N, 4)),
      summary_row
    )
  }

  out <- do.call(rbind, results)
  rownames(out) <- NULL
  out
}
```

# Run broad scans

```{r run_scans}
set.seed(42)

n_rep <- 50
factor_corr <- 0.3

# J-scan: fixed N, varying J.
result_J_df <- run_scan(
  scan_type = "J",
  J_values = seq(9, 45, by = 3),
  fixed_N = 300,
  factor_corr = factor_corr,
  n_rep = n_rep
)

# N-scan: fixed J, varying N.
result_N_df <- run_scan(
  scan_type = "N",
  N_values = seq(100, 1000, by = 100),
  fixed_J = 30,
  factor_corr = factor_corr,
  n_rep = n_rep
)

print(result_J_df, digits = 4)
print(result_N_df, digits = 4)
```

# Save result tables

```{r save_tables}
save_dir <- file.path("results", "tables", "broad_scans")
dir.create(save_dir, recursive = TRUE, showWarnings = FALSE)

write.csv(result_J_df, file.path(save_dir, "scan_J_corr03.csv"), row.names = FALSE)
write.csv(result_N_df, file.path(save_dir, "scan_N_corr03.csv"), row.names = FALSE)

cat("Saved result tables to:", save_dir, "\n")
```

# Plot function

```{r plot_function}
# Plot the 2x2 panel used in the thesis.
plot_broad_scan_2x2 <- function(df, x_var, x_label, title_text) {
  x <- df[[x_var]]

  # Keep model colors consistent across all thesis figures.
  col_lslx <- "steelblue"
  col_esem <- "red"

  par(mfrow = c(2, 2), mar = c(5, 5, 4, 2), oma = c(0, 0, 2, 0))

  # (a) MSE
  ylim <- range(c(
    df$lslx_mse_mean - df$lslx_mse_sd,
    df$lslx_mse_mean + df$lslx_mse_sd,
    df$esem_mse_mean - df$esem_mse_sd,
    df$esem_mse_mean + df$esem_mse_sd
  ), na.rm = TRUE)

  plot(x, df$lslx_mse_mean,
       type = "b", pch = 18, col = col_lslx, lty = 2, lwd = 2,
       ylim = ylim, xlab = x_label, ylab = "MSE",
       main = paste0("(a) MSE vs ", x_label), cex.main = 1.1)

  polygon(c(x, rev(x)),
          c(df$lslx_mse_mean + df$lslx_mse_sd,
            rev(df$lslx_mse_mean - df$lslx_mse_sd)),
          col = adjustcolor(col_lslx, 0.15), border = NA)

  lines(x, df$esem_mse_mean,
        type = "b", pch = 17, col = col_esem, lty = 2, lwd = 2)

  polygon(c(x, rev(x)),
          c(df$esem_mse_mean + df$esem_mse_sd,
            rev(df$esem_mse_mean - df$esem_mse_sd)),
          col = adjustcolor(col_esem, 0.15), border = NA)

  legend("topright", legend = c("lslx", "CC-ESEM"),
         col = c(col_lslx, col_esem), lty = 2, pch = c(18, 17),
         lwd = 2, bty = "n", cex = 0.9)

  # (b) Average absolute bias
  ylim <- range(c(df$lslx_ab, df$esem_ab), na.rm = TRUE)
  ylim <- ylim + c(-0.02, 0.08) * diff(ylim)

  plot(x, df$lslx_ab,
       type = "b", pch = 18, col = col_lslx, lty = 2, lwd = 2,
       ylim = ylim, xlab = x_label, ylab = "Average Absolute Bias",
       main = paste0("(b) AB vs ", x_label), cex.main = 1.1)

  lines(x, df$esem_ab,
        type = "b", pch = 17, col = col_esem, lty = 2, lwd = 2)

  legend("topright", legend = c("lslx", "CC-ESEM"),
         col = c(col_lslx, col_esem), lty = 2, pch = c(18, 17),
         lwd = 2, bty = "n", cex = 0.9)

  # (c) Variance
  ylim <- range(c(df$lslx_var, df$esem_var), na.rm = TRUE)
  ylim <- ylim + c(-0.02, 0.08) * diff(ylim)

  plot(x, df$lslx_var,
       type = "b", pch = 18, col = col_lslx, lty = 2, lwd = 2,
       ylim = ylim, xlab = x_label, ylab = "Average Variance",
       main = paste0("(c) VAR vs ", x_label), cex.main = 1.1)

  lines(x, df$esem_var,
        type = "b", pch = 17, col = col_esem, lty = 2, lwd = 2)

  legend("topright", legend = c("lslx", "CC-ESEM"),
         col = c(col_lslx, col_esem), lty = 2, pch = c(18, 17),
         lwd = 2, bty = "n", cex = 0.9)

  # (d) Runtime
  ylim <- c(0, max(df$lslx_time, df$esem_time, na.rm = TRUE) * 1.1)

  plot(x, df$lslx_time,
       type = "b", pch = 18, col = col_lslx, lty = 2, lwd = 2,
       ylim = ylim, xlab = x_label, ylab = "Runtime (sec)",
       main = paste0("(d) Runtime vs ", x_label), cex.main = 1.1)

  lines(x, df$esem_time,
        type = "b", pch = 17, col = col_esem, lty = 2, lwd = 2)

  legend("topleft", legend = c("lslx", "CC-ESEM"),
         col = c(col_lslx, col_esem), lty = 2, pch = c(18, 17),
         lwd = 2, bty = "n", cex = 0.9)

  mtext(title_text, outer = TRUE, cex = 1.1, font = 2)
  par(mfrow = c(1, 1))
}
```


# J-scan figure

```{r plot_j_scan, fig.width=9, fig.height=7}
plot_broad_scan_2x2(
  df = result_J_df,
  x_var = "J",
  x_label = "J",
  title_text = "J-scan Summary (factor_corr = 0.3, fixed N = 300, 50 reps)"
)
```

# N-scan figure

```{r plot_n_scan, fig.width=9, fig.height=7}
plot_broad_scan_2x2(
  df = result_N_df,
  x_var = "N",
  x_label = "N",
  title_text = "N-scan Summary (factor_corr = 0.3, fixed J = 30, 50 reps)"
)
```

