---
title: "Turning-point interval regression: rho = 0"
output:
  html_document:
    toc: true
    toc_depth: 2
---

# Purpose

This file reproduces the interval regression for the turning-point analysis under the orthogonal factor condition, `rho = 0`.

The turning intervals are manually recorded from the local refinement scans.  
This file does **not** rerun the simulations. It only fits the interval regressions and produces the regression figure.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Manually recorded turning intervals

```{r data}
interval_df <- data.frame(
  J  = c(15, 18, 21, 24, 27, 30, 33, 36, 39,
         42, 45, 48, 51, 54, 57, 60),

  # Lower and upper bounds of the manually identified turning J/N interval.
  # These values follow the manually recorded intervals used for the thesis figure/table.
  lo = c(15/130, 18/160, 21/180, 24/220, 27/240, 30/300,
         33/300, 36/300, 39/350, 42/400, 45/400, 48/400,
         51/500, 54/500, 57/500, 60/550),

  hi = c(15/120, 18/140, 21/160, 24/160, 27/210, 30/200,
         33/200, 36/250, 39/250, 42/350, 45/350, 48/350,
         51/400, 54/400, 57/400, 60/450)
)

interval_df$mid <- (interval_df$lo + interval_df$hi) / 2

print(interval_df, row.names = FALSE, digits = 5)

# Diagnostic only. It does not change the data.
if (any(interval_df$lo > interval_df$hi)) {
  warning("Some rows have lo > hi. This file keeps the original manually recorded values to reproduce the thesis results.")
}
```

# Fit interval regressions

```{r fit}
fit_mid <- lm(mid ~ J, data = interval_df)
fit_lo  <- lm(lo  ~ J, data = interval_df)
fit_hi  <- lm(hi  ~ J, data = interval_df)

cat("=== Midpoint Regression ===\n")
print(summary(fit_mid))

cat("\n=== Lower Bound Regression ===\n")
print(summary(fit_lo))

cat("\n=== Upper Bound Regression ===\n")
print(summary(fit_hi))

regression_summary <- data.frame(
  bound = c("midpoint", "lower", "upper"),
  intercept = c(coef(fit_mid)[1], coef(fit_lo)[1], coef(fit_hi)[1]),
  slope = c(coef(fit_mid)[2], coef(fit_lo)[2], coef(fit_hi)[2]),
  r_squared = c(summary(fit_mid)$r.squared,
                summary(fit_lo)$r.squared,
                summary(fit_hi)$r.squared),
  adj_r_squared = c(summary(fit_mid)$adj.r.squared,
                    summary(fit_lo)$adj.r.squared,
                    summary(fit_hi)$adj.r.squared)
)

print(regression_summary, row.names = FALSE, digits = 5)
```

# Plot

```{r plot, fig.width=8, fig.height=5}
plot_turning_regression <- function(interval_df, fit_mid, fit_lo, fit_hi) {
  par(mar = c(4, 4, 3, 1))

  J_seq <- seq(14, 61, length.out = 200)

  pred_mid <- predict(fit_mid, newdata = data.frame(J = J_seq), interval = "confidence")
  pred_lo  <- predict(fit_lo,  newdata = data.frame(J = J_seq))
  pred_hi  <- predict(fit_hi,  newdata = data.frame(J = J_seq))

  plot(NULL,
       xlim = c(14, 61),
       ylim = c(0, max(interval_df$hi) * 1.2),
       xlab = "J",
       ylab = "Estimated turning J/N",
       main = "Interval Regression: Turning J/N vs J (factor_corr = 0)")

  # midpoint confidence band
  polygon(c(J_seq, rev(J_seq)),
          c(pred_mid[, "lwr"], rev(pred_mid[, "upr"])),
          col = adjustcolor("#3266AD", 0.15),
          border = NA)

  # lower and upper bound regression lines
  lines(J_seq, pred_lo, col = "#D85A32", lwd = 1.5, lty = 2)
  lines(J_seq, pred_hi, col = "#D85A32", lwd = 1.5, lty = 2)

  # midpoint regression line
  lines(J_seq, pred_mid[, "fit"], col = "#3266AD", lwd = 2)

  # manually selected turning intervals
  arrows(x0 = interval_df$J, y0 = interval_df$lo,
         x1 = interval_df$J, y1 = interval_df$hi,
         angle = 90, code = 3, length = 0.04,
         col = "#D85A32", lwd = 1.5)

  # midpoint points
  points(interval_df$J, interval_df$mid,
         pch = 16, col = "#3266AD", cex = 1.3)

  # J labels
  text(interval_df$J, interval_df$hi,
       labels = paste0("J=", interval_df$J),
       pos = 3, cex = 0.7, col = "gray40")

  legend("bottomright",
         bty = "n",
         cex = 0.8,
         legend = c(
           paste0("Midpoint slope = ",
                  round(coef(fit_mid)[2], 5),
                  ", R² = ",
                  round(summary(fit_mid)$r.squared, 3)),
           paste0("Upper bound slope = ",
                  round(coef(fit_hi)[2], 5),
                  ", R² = ",
                  round(summary(fit_hi)$r.squared, 3)),
           paste0("Lower bound slope = ",
                  round(coef(fit_lo)[2], 5),
                  ", R² = ",
                  round(summary(fit_lo)$r.squared, 3)),
           "Manual turning interval"
         ),
         col = c("#3266AD", "#D85A32", "#D85A32", "#D85A32"),
         lty = c(1, 2, 2, NA),
         pch = c(NA, NA, NA, 16),
         lwd = c(2, 1.5, 1.5, NA))
}

plot_turning_regression(interval_df, fit_mid, fit_lo, fit_hi)
```


